---
title: "High_Dose_Ethanol_Manuscript"
format: html
editor: visual
---

## High Dose Alcohol Manuscript

This code is for the manuscript titled "High-dose Ethanol-induced Immunosuppression Modulates Sex-Specific Disease Outcomes in a Murine Model of Multiple Sclerosis". The code is maintained on the Melamed Lab github at the following link:
https://github.com/melamedlab/HighDoseEtOH_EAE

## Data Load

```{r}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(readxl)
library(emmeans)
library(ggbeeswarm)
library(writexl)
library(Hmisc)
library(sciplot)
library(rstatix)
library(stringr)
library(phyloseq)
library(vegan)
library(patchwork)
#library(ANCOMBC)
#library(curatedMetagenomicData)

theme_set(theme_classic(base_size = 13))

standard_error <- function(x) sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))

colors_treatment <- c("control" = "#415CAA", "etoh" = "#F26522",
                      "Ctl" = "#415CAA", "EtOH" = "#F26522")
colors_sex <- c("male" = "#F26522", "female" = "#F6B284")
shapes_sex <- c("male" = 16, "female" = 17)
output <- "../output/"
```

## Figure 2: Blood Ethanol Concentration (BEC) and Liver Histology
### Figure 2A: BEC Graph
```{r, fig.width = 3, fig.height = 3}

bec <- read_csv("../data/BEC_data.csv") %>%
  mutate(
    BEC_plot = ifelse(BEC < 0, 0, BEC),
    treatment = factor(treatment, levels = c("control", "etoh")),
    sex = factor(sex, levels = c("male", "female"))
  )

p_bec_treatment <- ggplot(bec, aes(treatment, BEC_plot)) +
  geom_boxplot(aes(color = treatment), outlier.shape = NA, fill = NA, width = 0.6) +
  geom_jitter(aes(color = treatment, shape = sex),
              width = 0.15, size = 2, alpha = 0.7) +
  scale_color_manual(values = colors_treatment) +
  scale_shape_manual(values = shapes_sex) +
  labs(x = "Treatment", y = "BEC (mg/dL)") +
  stat_compare_means(method = "wilcox.test", label = "p.signif")
p_bec_treatment
ggsave(paste0(output, "Figure2A.pdf"), p_bec_treatment, width = 3, height = 3)
```
### Figure 2B and 2C: Rep Images
Figure 2B and 2C are representative images and have no corresponding code.

### Figure 2D: Oil O Red
```{r, fig.width = 3, fig.height = 3}

oro <- read_excel("../data/Oil_Red_O_data.xlsx") %>%
  mutate(
    Treatment = factor(Treatment, levels = c("Ctl", "EtOH")),
    Sex = factor(Sex, levels = c("male", "female"))
  )

p_oro_treatment <- ggplot(oro, aes(Treatment, ORO_value)) +
  geom_boxplot(aes(color = Treatment), outlier.shape = NA,
               fill = NA, width = 0.6) +
  geom_jitter(aes(color = Treatment, shape = Sex),
              width = 0.15, size = 2) +
  scale_color_manual(values = colors_treatment) +
  scale_shape_manual(values = shapes_sex) +
  labs(x = "Treatment", y = "Oil Red O lipid area") +
  stat_compare_means(method = "wilcox.test", label = "p.signif")
p_oro_treatment
ggsave("../output/Figure2D.pdf", width = 3, height = 3)

oro_summary <- oro %>%
  group_by(Treatment) %>%
  summarise(
    mean = mean(ORO_value),
    median = median(ORO_value),
    sd = sd(ORO_value),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    fold_change_vs_Ctl = ifelse(
      Treatment == "EtOH",
      mean / mean[Treatment == "Ctl"],
      NA
    )
  )

write_csv(oro_summary, "../output/Figure2D_summary_statistics.csv")

```

## Figure 3: EAE Score Curves
### Figure 3A-E: EAE Score Curves
```{r}
group_colors <- c(
  "Ctl-Female"  = "#9BB9E1",
  "Ctl-Male"    = "#415CAA",
  
  "EtOH-Female" = "#F6B284",
  "EtOH-Male"   = "#F26522",
  
  "Ctl"         = "#7A7A7A",  
  "EtOH"        = "#000000"
)

standard_error <- function(x) {
  sd(x, na.rm = TRUE) / sqrt(length(na.omit(x)))
}

t_test_p_value <- function(col_A, col_B) {
  t.test(col_A, col_B)$p.value
}

eae_data <- read_csv("../data/EAE_data.csv")

eae_data <- eae_data %>%
  mutate(pattern = paste(Treatment, Sex, sep = "-"))

comparison.plot <- function(pattern1, pattern2, figure_name) {
  
  stopifnot(pattern1 %in% names(group_colors),
            pattern2 %in% names(group_colors))
  
  col1 <- group_colors[pattern1]
  col2 <- group_colors[pattern2]
  
  gr1 <- eae_data %>%
    filter(str_detect(pattern, pattern1)) %>%
    select(`7`:`19`)
  
  gr2 <- eae_data %>%
    filter(str_detect(pattern, pattern2)) %>%
    select(`7`:`19`)
  
  gr1.means <- summarise(gr1, across(everything(), mean, na.rm = TRUE)) %>% unlist()
  gr1.stdev <- summarise(gr1, across(everything(), standard_error)) %>% unlist()
  
  gr2.means <- summarise(gr2, across(everything(), mean, na.rm = TRUE)) %>% unlist()
  gr2.stdev <- summarise(gr2, across(everything(), standard_error)) %>% unlist()
  
  sigs <- mapply(t_test_p_value, gr1, gr2)
  days <- as.numeric(colnames(gr1))
  
  pdf(paste(output, figure_name, ".pdf"),
      width = 6, height = 4)
  
  plot(days, gr1.means, type = "o",
       col = col1, frame.plot = FALSE, axes = FALSE,
       xlim = c(7, 19), ylim = c(0, 4),
       xlab = "Day", ylab = "Score")
  
  axis(2, at = 0:4)
  axis(1, at = 7:19)
  
  lines(days, gr2.means, type = "o", col = col2)
  
  errbar(days, gr1.means,
         gr1.means + gr1.stdev,
         gr1.means - gr1.stdev,
         add = TRUE, errbar.col = col1, col = col1)
  
  errbar(days, gr2.means,
         gr2.means + gr2.stdev,
         gr2.means + (-gr2.stdev),
         add = TRUE, errbar.col = col2, col = col2)
  
  sel <- (0.01 < sigs & sigs < 0.05)
  text(days[sel], pmax(gr1.means[sel], gr2.means[sel]) + 0.45, "*", cex = 1.5)
  
  sel <- (0.001 < sigs & sigs < 0.01)
  text(days[sel], pmax(gr1.means[sel], gr2.means[sel]) + 0.45, "**", cex = 1.5)
  
  sel <- sigs < 0.001
  text(days[sel], pmax(gr1.means[sel], gr2.means[sel]) + 0.45, "***", cex = 1.5)
  
  legend("topleft",
         legend = c(pattern1, pattern2),
         col = c(col1, col2),
         lty = 1,
         pch = 16,
         bty = "n")
  
  dev.off()
}

comparison.plot("EtOH", "Ctl", "Figure3A")
comparison.plot("EtOH-Male", "Ctl-Male", "Figure3B")
comparison.plot("EtOH-Female", "Ctl-Female", "Figure3C")
comparison.plot("Ctl-Male", "Ctl-Female", "Figure3D")
comparison.plot("EtOH-Male", "EtOH-Female", "Figure3E")
```

### Figure 3F: Days >= EAE Score of 2
```{r, fig.width = 4, fig.height = 3}

eae_data <- read_csv("../data/EAE_data.csv") %>%
  mutate(
    Treatment = factor(Treatment, levels = c("Ctl", "EtOH")),
    Sex = factor(Sex, levels = c("Female", "Male"))
  )

peak_days <- as.character(11:17)

eae_data <- eae_data %>%
  rowwise() %>%
  mutate(high_EAE_days = sum(c_across(all_of(peak_days)) >= 2, na.rm = TRUE)) %>%
  ungroup()

poisson_model <- glm(high_EAE_days ~ Treatment * Sex,
                     family = poisson, data = eae_data)

emm <- emmeans(poisson_model, ~ Treatment | Sex, type = "response")

contrast_df <- contrast(emm, "revpairwise", adjust = "fdr") %>%
  as.data.frame() %>%
  mutate(
    fold_change = round(ratio, 2),
    percent_change = round((1 - fold_change) * 100),
    signif = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      TRUE ~ "ns"
    )
  )

write_xlsx(contrast_df, "../output/Figure3F_stats.xlsx")

p_eae <- ggplot(eae_data,
                aes(Treatment, high_EAE_days, color = Treatment)) +
  geom_boxplot(width = 0.55, outlier.shape = NA, alpha = 0.25) +
  geom_quasirandom(width = 0.15, size = 2, alpha = 0.7) +
  scale_color_manual(values = colors_treatment) +
  facet_wrap(~ Sex) +
  labs(
    x = "Treatment",
    y = expression("Number of days with EAE score " >= 2)
  ) +
  theme(legend.position = "none")
p_eae
ggsave("../output/Figure3F.pdf", p_eae, width = 4, height = 3)

```


## Figure 4: Microbiome
### Microbiome Data Load
```{r}
group_colors <- c(
  "EtOH Male"   = "#F26522",
  "EtOH Female" = "#F6B284",
  "Ctl Male"    = "#415CAA",
  "Ctl Female"  = "#9BB9E1"
)

group_levels <- c("EtOH Male", "EtOH Female", "Ctl Male", "Ctl Female")

metadata <- read_xlsx("../data/MBX6_metadata.xlsx") %>%
  mutate(
    sample_id = gsub("-", "_", sample_id),
    sex = factor(sex),
    GroupSex = factor(paste(group, sex), levels = group_levels),
    timepoint = factor(adriana_timepoint,
                       levels = c("pre-EAE (D0)", "Day 26 (D26)"),
                       labels = c("D0", "D26"))
  )
setwd("../data/microbiome_data/")
profile_files <- list.files(pattern = "_taxonomic_profile\\.tsv$", full.names = TRUE)

read_profile <- function(file) {
  read_tsv(file, comment = "#",
           col_names = c("clade", "tax_id", "abundance", "extra"),
           col_types = cols(.default = "c")) %>%
    mutate(
      abundance = as.numeric(abundance),
      sample_id = basename(file) %>%
        str_remove("_taxonomic_profile\\.tsv") %>%
        str_replace("pre-EAE", "-D0") %>%
        str_replace("post-EAE", "-D26")
    ) %>%
    filter(!is.na(abundance)) %>%
    select(taxon = clade, abundance, sample_id)
}

abundance_long <- map_dfr(profile_files, read_profile)
```
### Figure 4A: Shannon Diversity
```{r, fig.width = 5, fig.height = 4}
group_colors <- c(
  "Ctl Female"  = "#9BB9E1",
  "Ctl Male"    = "#415CAA",
  "EtOH Female" = "#F6B284",
  "EtOH Male"   = "#F26522"
)
shannon_df_treatment <- read_csv("../data/shannon_abundance.csv") %>%
  filter(Metric == "Shannon") %>%
   left_join(metadata %>% mutate(sample_id = gsub("_", "-", sample_id)),
             by = c("SampleID" = "sample_id"))

shannon_df_treatment <- shannon_df_treatment %>%
  mutate(
    Group = factor(Group, levels = c("Ctl Female", "Ctl Male", "EtOH Female", "EtOH Male")),
    timepoint = factor(timepoint)
  )

pairwise_stats <- shannon_df_treatment %>%
  group_by(timepoint) %>%
  wilcox_test(Value ~ Group, p.adjust.method = "BH") %>%
  add_significance("p.adj") %>%
  filter(p.adj < 0.05) %>%                       # keep only significant comparisons
  add_xy_position(x = "Group", fun = "max")   # automatic y positions


 p <- ggplot(shannon_df_treatment, aes(x = Group, y = Value, color = Group)) +
    geom_boxplot(width = 0.6, fill = NA, outlier.shape = NA) +  # transparent boxes
    geom_jitter(width = 0.15, size = 2, alpha = 0.7) +                    # colored dots
    facet_wrap(~timepoint) +
    scale_color_manual(values = group_colors) +
    stat_pvalue_manual(
      pairwise_stats,
      label = "p.adj.signif",
      tip.length = 0.02) +
    theme_classic(base_size = 12) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none") +
    labs(x = "Treatment × Sex", y = "Shannon Diversity")
p
ggsave("../output/Figure4A.pdf", width = 4, height = 5)

```

### Figure 4B: Bray-Curtis PCoA
```{r, fig.width = 8, fig.height = 4}
otu_mat <- abundance_long %>%
  pivot_wider(names_from = sample_id, values_from = abundance, values_fill = 0) %>%
  column_to_rownames("taxon") %>%
  as.matrix() %>%
  {.[rowSums(.) > 0, colSums(.) > 0]} # remove empty rows/columns

ps <- phyloseq(
  otu_table(otu_mat, taxa_are_rows = TRUE),
  sample_data(metadata %>% mutate(sample_id = gsub("_", "-", sample_id)) %>%
                filter(sample_id %in% colnames(otu_mat)) %>%
                column_to_rownames("sample_id"))
)

ps_D0  <- prune_samples(ps@sam_data[["timepoint"]] == "D0", ps)
ps_D26 <- prune_samples(ps@sam_data[["timepoint"]] == "D26", ps)

ord_D0  <- ordinate(ps_D0,  "PCoA", "bray")
ord_D26 <- ordinate(ps_D26, "PCoA", "bray")

axis_labels <- function(ord) {
  paste0("PC", 1:2, " (", round(100 * ord$values$Relative_eig[1:2], 1), "%)")
}

plot_pcoa <- function(ps_obj, ord, title) {
  plot_ordination(ps_obj, ord, color = "GroupSex") +
    geom_point(size = 3) +
    scale_color_manual(values = group_colors) +
    labs(
      title = title,
      x = axis_labels(ord)[1],
      y = axis_labels(ord)[2],
      color = "Treatment × Sex"
    ) +
    theme_bw(base_size = 14)
}

pcoa_plot <- ggpubr::ggarrange(
  plot_pcoa(ps_D0, ord_D0, "PCoA – D0"),
  plot_pcoa(ps_D26, ord_D26, "PCoA – D26"),
  ncol = 2, common.legend = T
)
pcoa_plot
ggsave("../output/Figure4B.pdf", width = 12, height = 6)
```

### Figure 4C: Phlyum Barcharts
```{r, fig.width = 10, fig.height = 5}
c25 <- c(
  "dodgerblue2","#E31A1C","green4","#6A3D9A","#FF7F00",
  "black","gold1","skyblue2","#FB9A99","palegreen2",
  "#CAB2D6","#FDBF6F","gray70","khaki2","maroon",
  "orchid1","deeppink1","blue1","steelblue4","darkturquoise",
  "green1","yellow4","yellow3","darkorange4","brown"
)

phylum_data <- abundance_long %>%
  dplyr::rename("bacteria" = "taxon") %>%
  filter(str_detect(bacteria, "p__") & !str_detect(bacteria, "c__")) %>%
  left_join(metadata %>% mutate(sample_id = gsub("_", "-", sample_id)), by = "sample_id") %>%
  group_by(group, sex, adriana_timepoint, bacteria) %>%
  summarise(mean_abundance = mean(abundance), .groups = "drop") %>%
  mutate(
    bacteria = gsub(".*p__", "", bacteria),
    TGS = paste(group, sex)
  )

top_phyla <- phylum_data %>%
  group_by(bacteria) %>%
  summarise(total = sum(mean_abundance), .groups = "drop") %>%
  slice_max(total, n = 10) %>%
  pull(bacteria)

phylum_plot <- phylum_data %>%
  filter(bacteria %in% top_phyla) %>%
  mutate(adriana_timepoint = factor(adriana_timepoint, c("pre-EAE (D0)", "Day 26 (D26)"))) %>%
  ggplot(aes(TGS, mean_abundance, fill = bacteria)) +
  geom_bar(stat = "identity") +
  facet_grid(~adriana_timepoint) +
  theme_classic() +
  ylab("Mean Relative Abundance (%)") +
  xlab("Treatment × Sex") +
  scale_fill_manual(values = c25)
phylum_plot
ggsave("../output/Figure4C.pdf", width = 10, height = 5)

```

### Figure 4D: Genus Barcharts
```{r, fig.width = 10, fig.height = 5}
genus_tmp <- abundance_long %>%
  dplyr::rename("bacteria" = "taxon") %>%
  filter(str_detect(bacteria, "g__") & !str_detect(bacteria, "s__")) %>%
  left_join(metadata %>% mutate(sample_id = gsub("_", "-", sample_id)), by = "sample_id") %>%
  group_by(group, sex, adriana_timepoint, bacteria) %>%
  summarise(mean = mean(abundance), .groups = "drop") %>%
  mutate(
    bacteria = gsub(".*g__", "", bacteria),
    TGS = paste(group, sex)
  )

top_genera <- genus_tmp %>%
  group_by(bacteria) %>%
  summarise(total = sum(mean), .groups = "drop") %>%
  slice_max(total, n = 10) %>%
  pull(bacteria)

plot_data <- genus_tmp %>%
  filter(bacteria %in% top_genera) %>%
  group_by(TGS, adriana_timepoint) %>%
  mutate(mean_norm = mean / sum(mean) * 100) %>%
  ungroup() %>%
   mutate(adriana_timepoint = factor(adriana_timepoint, c("pre-EAE (D0)", "Day 26 (D26)")))

genus_plot <- ggplot(plot_data, aes(TGS, mean_norm, fill = bacteria)) +
  geom_bar(stat = "identity") +
  facet_grid(~adriana_timepoint) +
  theme_classic() +
  ylab("Relative Abundance (%)") +
  xlab("Treatment × Sex") +
  scale_fill_manual(values = c25)
genus_plot
ggsave("../output/Figure4D.pdf", width = 10, height = 5)
```

### Figure 4E: FB Ratio
```{r, fig.width = 5, fig.height = 4}
group_colors <- c(
  "Ctl Female"  = "#9BB9E1",
  "Ctl Male"    = "#415CAA",
  "EtOH Female" = "#F6B284",
  "EtOH Male"   = "#F26522"
)

phylum_data <- abundance_long %>%
  mutate(taxon = gsub("k__Bacteria\\|", "", taxon)) %>%
  filter(str_detect(taxon, "^p__")) %>%
  mutate(phylum = str_extract(taxon, "(?<=p__)[^|]+")) %>%
  filter(!grepl("\\|c__", taxon)) %>%
  left_join(metadata %>% mutate(sample_id = gsub("_", "-", sample_id)), by = "sample_id")

ratio_data <- phylum_data %>%
  filter(phylum %in% c("Firmicutes", "Bacteroidota", "Verrucomicrobia")) %>%
  select(sample_id, GroupSex, timepoint, phylum, abundance) %>%
  pivot_wider(names_from = phylum, values_from = abundance, values_fill = 0) %>%
  mutate(
    FB_ratio = ifelse(Bacteroidota > 0, Firmicutes / Bacteroidota, NA),
    VF_ratio = ifelse(Firmicutes > 0, Verrucomicrobia / Firmicutes, NA)
  ) %>%
  pivot_longer(cols = c(FB_ratio, VF_ratio), names_to = "ratio", values_to = "value")

fb_stats <- ratio_data %>%
   mutate(GroupSex = factor(GroupSex, levels = c("Ctl Female", "Ctl Male",
                                                           "EtOH Female", "EtOH Male"))) %>%
  filter(ratio == "FB_ratio") %>%
  group_by(timepoint) %>%
  pairwise_wilcox_test(value ~ GroupSex, p.adjust.method = "BH") %>%
  add_significance("p.adj") %>%
  filter(p.adj < 0.05) %>%
  add_xy_position(step.increase = 0.01)


gg4E <- ratio_data %>%
             mutate(GroupSex = factor(GroupSex, levels = c("Ctl Female", "Ctl Male",
                                                           "EtOH Female", "EtOH Male"))) %>%
             filter(ratio == "FB_ratio") %>%
    ggplot(aes(x = GroupSex, y = value, color = GroupSex)) +
    geom_boxplot(width = 0.6, outlier.shape = NA) +
    geom_jitter(width = 0.3, size = 1.5, alpha = 0.7) +
    scale_color_manual(values = group_colors) +
    stat_pvalue_manual(
      fb_stats,
      label = "p.adj.signif"
    ) +
    facet_wrap(~ timepoint, scales = "free_y") +
    theme_classic(base_size = 14) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none"
    ) +
    labs(x = "Treatment × Sex", y = "FB Ratio")
gg4E

ggsave("../output/Figure4E.pdf", width = 5, height = 4)
```

### Figure 4F: VF Ratio
```{r, fig.width = 5, fig.height = 4}
group_colors <- c(
  "Ctl Female"  = "#9BB9E1",
  "Ctl Male"    = "#415CAA",
  "EtOH Female" = "#F6B284",
  "EtOH Male"   = "#F26522"
)

phylum_data <- abundance_long %>%
  mutate(taxon = gsub("k__Bacteria\\|", "", taxon)) %>%
  filter(str_detect(taxon, "^p__")) %>%
  mutate(phylum = str_extract(taxon, "(?<=p__)[^|]+")) %>%
  filter(!grepl("\\|c__", taxon)) %>%
  left_join(metadata %>% mutate(sample_id = gsub("_", "-", sample_id)), by = "sample_id")

ratio_data <- phylum_data %>%
  filter(phylum %in% c("Firmicutes", "Bacteroidota", "Verrucomicrobia")) %>%
  select(sample_id, GroupSex, timepoint, phylum, abundance) %>%
  pivot_wider(names_from = phylum, values_from = abundance, values_fill = 0) %>%
  mutate(
    FB_ratio = ifelse(Bacteroidota > 0, Firmicutes / Bacteroidota, NA),
    VF_ratio = ifelse(Firmicutes > 0, Verrucomicrobia / Firmicutes, NA)
  ) %>%
  pivot_longer(cols = c(FB_ratio, VF_ratio), names_to = "ratio", values_to = "value")

vf_stats <- ratio_data %>%
   mutate(GroupSex = factor(GroupSex, levels = c("Ctl Female", "Ctl Male",
                                                           "EtOH Female", "EtOH Male"))) %>%
  filter(ratio == "VF_ratio") %>%
  group_by(timepoint) %>%
  pairwise_wilcox_test(value ~ GroupSex, p.adjust.method = "BH") %>%
  add_significance("p.adj") %>%
  filter(p.adj < 0.05) %>%
  add_xy_position(step.increase = 0.1)

gg4F <- ratio_data %>%
             mutate(GroupSex = factor(GroupSex, levels = c("Ctl Female", "Ctl Male",
                                                           "EtOH Female", "EtOH Male"))) %>%
             filter(ratio == "VF_ratio") %>%
    ggplot(aes(x = GroupSex, y = value, color = GroupSex)) +
    geom_boxplot(width = 0.6, outlier.shape = NA) +
    geom_jitter(width = 0.3, size = 1.5, alpha = 0.7) +
    scale_color_manual(values = group_colors) +
    stat_pvalue_manual(
      vf_stats,
      label = "p.adj.signif"
    ) +
    facet_wrap(~ timepoint, scales = "free_y") +
    theme_classic(base_size = 14) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none"
    ) +
    labs(x = "Treatment × Sex", y = "VF Ratio")
gg4F
ggsave("../output/Figure4F.pdf", width = 5, height = 4)
```

## Figure 5: Genus and Phylum Changes 
### Figure 5A-D: Phylum Changes
```{r, fig.width = 5, fig.height = 7}
out_df <- data.frame()
colors <- c(
  "Ctl Female"  = "#9BB9E1",
  "Ctl Male"    = "#415CAA",
  "EtOH Female" = "#F6B284",
  "EtOH Male"   = "#F26522"
)
active_level = "phylum"
taxa_level <- abundance_long %>%
  complete(taxon, sample_id) %>%
  mutate(abundance = ifelse(is.na(abundance), 0, abundance)) %>%
  dplyr::rename("bacteria" = "taxon",
                "relative_abundance" = "abundance") %>%
  filter(grepl("p__", bacteria)) %>%
  filter(!grepl("c__", bacteria)) %>%
  left_join(metadata %>% mutate(sample_id = gsub("_", "-", sample_id))) %>%
  dplyr::mutate(mean = (relative_abundance)) %>%
  mutate(bacteria = gsub(".*p__", "", bacteria)) %>%
  mutate(TGS = paste0(group, " ", sex)) %>%
  mutate(adriana_timepoint = factor(adriana_timepoint, c("pre-EAE (D0)",
                                                         "Day 26 (D26)")))

check <- taxa_level %>%
  left_join(metadata %>% mutate(sample_id = gsub("_", "-", sample_id))) %>%
  as.data.frame() %>%
  compare_means(data = ., mean ~ TGS, group.by = c("bacteria", "timepoint")) %>%
  select(-p.adj) %>%
  filter(timepoint == "D26") %>%
  group_by(bacteria) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj") %>%
  mutate(taxa_level = "phylum")
write.csv(check, "../output/phylum_statistical_results_D26.csv")

## In the manuscript we focus on the phyla: Bacteroidota, Verrucomicrobia, Actinobacteria, and Firmicutes
## in the main figure. This is due to their large contribution and known
## physiological roles.
check <- check %>%
  filter(bacteria %in% c("Verrucomicrobia", "Bacteroidota",
                         "Actinobacteria", "Firmicutes"))

plot.list <- list()
for(active_taxa in unique(check$bacteria[check$p.adj <= 0.05])){
  tmp.data <- taxa_level %>%
    left_join(metadata) %>%
    filter(bacteria == active_taxa) %>%
    filter(grepl("D26", timepoint))
  
  max_value <- max(tmp.data$mean)
 tmpgg <- tmp.data %>%
         ggplot(aes(x=TGS, y=mean)) +
         geom_boxplot(outlier.shape = NA, aes(color = TGS)) +
         geom_quasirandom(aes(color = TGS)) +
         ggtitle(active_taxa) +
         ylab("Relative Abundance (%)") + 
         stat_pvalue_manual(check %>%
                              filter(bacteria == active_taxa) %>%
                              filter(p.adj <= 0.05) %>%
                              mutate(y.position =rep_len(max_value * c(1.07, 1.14, 1.21, 1.28, 1.35, 1.42),
                                                         length.out = nrow(.))), label = "p.adj.signif") +
         scale_color_manual(values = colors) +
        theme_bw() +
         theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
                     axis.text = element_text(color = "black"),
                     legend.position = "none",
                     axis.title.x = element_blank()) 
 
 plot.list <- c(plot.list, list(tmpgg))
}

ggarrange(plot.list[[4]], plot.list[[2]], plot.list[[1]], plot.list[[3]], 
          nrow = 2, ncol = 2, common.legend = T)
ggsave("../output/Figure5A_D.pdf", width = 5, height = 7)
```

### Figure 5E-J: Genius Changes
```{r, fig.width = 7.5, fig.height = 7}
colors <- c(
  "Ctl Female"  = "#9BB9E1",
  "Ctl Male"    = "#415CAA",
  "EtOH Female" = "#F6B284",
  "EtOH Male"   = "#F26522"
)
active_level = "genius"
taxa_level <- abundance_long %>%
  complete(taxon, sample_id) %>%
  mutate(abundance = ifelse(is.na(abundance), 0, abundance)) %>%
  dplyr::rename("bacteria" = "taxon",
                "relative_abundance" = "abundance") %>%
  filter(grepl("g__", bacteria)) %>%
  filter(!grepl("s__", bacteria)) %>%
  left_join(metadata %>% mutate(sample_id = gsub("_", "-", sample_id))) %>%
  dplyr::mutate(mean = (relative_abundance)) %>%
  mutate(bacteria = gsub(".*g__", "", bacteria)) %>%
  mutate(TGS = paste0(group, " ", sex)) %>%
  mutate(adriana_timepoint = factor(adriana_timepoint, c("pre-EAE (D0)",
                                                         "Day 26 (D26)")))

check <- taxa_level %>%
  left_join(metadata %>% mutate(sample_id = gsub("_", "-", sample_id))) %>%
  as.data.frame() %>%
  compare_means(data = ., mean ~ TGS, group.by = c("bacteria", "timepoint")) %>%
  select(-p.adj) %>%
  filter(timepoint == "D26") %>%
  group_by(bacteria) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj") %>%
  mutate(taxa_level = "genius")
write.csv(check, "../output/genius_statistical_results_D26.csv")

## In the manuscript we focus on the genera: Akkermansia, Bacteroides, Alistipes,
## GGB14001, GGB24132, GGB29685
check <- check %>%
  filter(bacteria %in% c("Akkermansia", "Bacteroides",
                         "Alistipes", "GGB14001",
                         "GGB24132", "GGB29685"))

plot.list <- list()
for(active_taxa in unique(check$bacteria[check$p.adj <= 0.05])){
  tmp.data <- taxa_level %>%
    left_join(metadata) %>%
    filter(bacteria == active_taxa) %>%
    filter(grepl("D26", timepoint))
  
  max_value <- max(tmp.data$mean)
 tmpgg <- tmp.data %>%
         ggplot(aes(x=TGS, y=mean)) +
         geom_boxplot(outlier.shape = NA, aes(color = TGS)) +
         geom_quasirandom(aes(color = TGS)) +
         ggtitle(active_taxa) +
         ylab("Relative Abundance (%)") + 
         stat_pvalue_manual(check %>%
                              filter(bacteria == active_taxa) %>%
                              filter(p.adj <= 0.05) %>%
                              mutate(y.position =rep_len(max_value * c(1.07, 1.14, 1.21, 1.28, 1.35, 1.42),
                                                         length.out = nrow(.))), label = "p.adj.signif") +
         scale_color_manual(values = colors) +
        theme_bw() +
         theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
                     axis.text = element_text(color = "black"),
                     legend.position = "none",
                     axis.title.x = element_blank()) 
 
 plot.list <- c(plot.list, list(tmpgg))
}

ggarrange(plot.list[[1]], plot.list[[3]], plot.list[[2]],
          plot.list[[4]], plot.list[[5]], plot.list[[6]], 
          nrow = 2, ncol = 3, common.legend = T)
ggsave("../output/Figure5E_J.pdf", width = 7.5, height = 7)
```

## Figure 6: Flow Cytometry

```{r}
color_pal <- c("#415CAA", "#F26522")
## Load the Flow Data, gating scheme for which is provided in Supplementary Figure 2
data <- read_xlsx("../data/Flow_data.xlsx") %>%
  dplyr::rename("mouse_id" = "...1")
colnames(data) <- gsub(" \\|.*", "", gsub(".*CD3\\+\\/", "",gsub(".*Live\\/", "",  colnames(data))))
colnames(data)[15] <- paste0(colnames(data)[15], " of CD69+ PD1+")

global_stats <- data %>%
  pivot_longer(cols = c(-mouse_id, -Group)) %>%
  group_by(name) %>%
  as.data.frame() %>%
  compare_means(data = ., value~Group, group.by = "name", method = "wilcox.test")

write.csv(global_stats, paste0(output, "Figure6_statistical_results.csv"))
```

### Figure 6A: B and T Cells
```{r, fig.width = 2.75, fig.height = 2.5}
gga <- data %>%
  pivot_longer(c(-`mouse_id`, -Group)) %>%
  filter(name %in% c("CD3+", "CD19+")) %>%
  ggplot(aes(x=Group, y= value, color = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.6) +
  ggbeeswarm::geom_quasirandom() +
  theme_classic() +
  facet_grid(~name) +
  ylab("Percent of\nLive Lymphocytes (%)") +
  stat_compare_means(comparisons = list(c("EtOH", "Ctl")),
                     method = "wilcox.test") +
  scale_color_manual(values = color_pal) +
  theme(axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        strip.background = element_blank(),
        legend.position = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))
gga + ylab("Percent of Live Lymphocytes (%)")
ggsave(filename = "Figure6A.pdf", path = "../output", width = 2.75, height = 2.5)
```

### Figure 6B: CD4+ and CD8+ T Cells
```{r, fig.width = 2.75, fig.height = 2.5}
ggb <- data %>%
  pivot_longer(c(-`mouse_id`, -Group)) %>%
  filter(name %in% c("CD4+", "CD8+")) %>%
  ggplot(aes(x=Group, y= value, color = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.6) +
  ggbeeswarm::geom_quasirandom() +
  theme_classic() +
  facet_grid(~name)  +
  ylab("Percent of T Cells (%)") +
  stat_compare_means(comparisons = list(c("EtOH", "Ctl")),
                     method = "wilcox.test") +
  scale_color_manual(values = color_pal)+
  theme(axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        strip.background = element_blank(),
        legend.position = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))
ggb
ggsave(filename = "Figure6B.pdf", path = "../output", width = 2.75, height = 2.5)
```

### Figure 6C: CD4+ Activation States
```{r, fig.width = 4.125, fig.height = 2.5}
ggc <- data %>%
  pivot_longer(c(-`mouse_id`, -Group)) %>%
  filter(name %in% c("CD4+/CD69+", "CD4+/OX40+",
                     "CD4+/PD-1+")) %>%
  ggplot(aes(x=Group, y= value, color = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.6) +
  ggbeeswarm::geom_quasirandom() +
  theme_classic() +
  facet_grid(~name) +
  ylab("Percent of CD4+ T cells (%)")  +
  stat_compare_means(comparisons = list(c("EtOH", "Ctl")),
                     method = "wilcox.test") +
  scale_color_manual(values = color_pal) +
  theme(axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        strip.background = element_blank(),
        legend.position = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) 
ggc
ggsave(filename = "Figure6C_part1.pdf", path = "../output", width = 4.125, height = 2.5)
```




```{r, fig.width = 5.5, fig.height = 2.5}
ggc2 <- data %>%
  pivot_longer(c(-`mouse_id`, -Group)) %>%
  filter(name %in% c("CD4+/Q1: CD69+ PD1-", "CD4+/Q2: CD69+ PD1+",
                     "CD4+/Q3: CD69- PD1+", "CD4+/Q4: CD69- PD1-")) %>%
  mutate(name = gsub(".*: ", "", name)) %>%
  mutate(name = factor(name, c("CD69- PD1-", "CD69+ PD1-",
                               "CD69+ PD1+", "CD69- PD1+"))) %>%
  ggplot(aes(x=Group, y= value, color = Group)) +
  geom_boxplot(outlier.shape = NA) +
  ggbeeswarm::geom_quasirandom() +
  theme_classic() +
  facet_wrap(~name, scales = "free", nrow = 1) +
  ylab("Percent of CD4+ T cells (%)")  +
  stat_compare_means(comparisons = list(c("EtOH", "Ctl")),
                     method = "wilcox.test") +
  scale_color_manual(values = color_pal)+
  theme(axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        strip.background = element_blank(),
        legend.position = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) 
ggc2
ggsave(filename = "Figure6C_part2.pdf", path = "../output", width = 5.5, height = 2.5)
```

### Figure 6D: CD8+ Activation States
```{r, fig.width = 2.75, fig.height = 2.5}
ggd <- data %>%
  pivot_longer(c(-`mouse_id`, -Group)) %>%
  filter(name %in% c("CD8+/CD69+", 
                     "CD8+/PD-1+")) %>%
  ggplot(aes(x=Group, y= value, color = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.6) +
  ggbeeswarm::geom_quasirandom() +
  theme_classic() +
  facet_grid(~name) +
  ylab("Percent of CD8+ T cells (%)")  +
  stat_compare_means(comparisons = list(c("EtOH", "Ctl")),
                     method = "wilcox.test") +
  scale_color_manual(values = color_pal) +
  theme(axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        strip.background = element_blank(),
        legend.position = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) 
ggd
ggsave(filename = "Figure6D_part1.pdf", path = "../output", width = 2.75, height = 2.5)
```

```{r, fig.width = 5.5, fig.height = 2.5}

ggd2 <- data %>%
  pivot_longer(c(-`mouse_id`, -Group)) %>%
  filter(name %in% c("CD8+/Q1: CD69+ PD1-", "CD8+/Q2: CD69+ PD1+",
                     "CD8+/Q3: CD69- PD1+", "CD8+/Q4: CD69- PD1-")) %>%
  mutate(name = gsub(".*: ", "", name)) %>%
  mutate(name = factor(name, c("CD69- PD1-", "CD69+ PD1-",
                               "CD69+ PD1+", "CD69- PD1+"))) %>%
  ggplot(aes(x=Group, y= value, color = Group)) +
  geom_boxplot(outlier.shape = NA) +
  ggbeeswarm::geom_quasirandom() +
  theme_classic() +
  facet_wrap(~name, scales = "free", nrow = 1) +
  ylab("Percent of CD8+ T cells (%)")  +
  stat_compare_means(comparisons = list(c("EtOH", "Ctl")),
                     method = "t.test") +
  scale_color_manual(values = color_pal)+
  theme(axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        strip.background = element_blank(),
        legend.position = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) 
ggd2
ggsave(filename = "Figure6D_part2.pdf", path = "../output", width = 5.5, height = 2.5)


```

### Figure 6E: DN Activation State
```{r, fig.width = 5.5, fig.height = 2.5}

gge <- data %>%
  pivot_longer(c(-`mouse_id`, -Group)) %>%
  filter(name %in% c("DN/Q1: CD69+ PD1-", "DN/Q2: CD69+ PD1+",
                     "DN/Q3: CD69- PD1+", "DN/Q4: CD69- PD1-")) %>%
  mutate(name = gsub(".*: ", "", name)) %>%
  mutate(name = factor(name, c("CD69- PD1-", "CD69+ PD1-",
                               "CD69+ PD1+", "CD69- PD1+"))) %>%
  ggplot(aes(x=Group, y= value, color = Group)) +
  geom_boxplot(outlier.shape = NA) +
  ggbeeswarm::geom_quasirandom() +
  theme_classic() +
  facet_wrap(~name, scales = "free", nrow = 1) +
  ylab("Percent of DN T cells (%)")  +
  stat_compare_means(comparisons = list(c("EtOH", "Ctl")),
                     method = "t.test") +
  scale_color_manual(values = color_pal)+
  theme(axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        strip.background = element_blank(),
        legend.position = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) 
gge
ggsave(filename = "Figure6E.pdf", path = "../output", width = 5.5, height = 2.5)
```


## Supplementary Figure 1
### Supplementary Figure 1A
```{r}
group_colors <- c(
  "Ctl-Female"  = "#9BB9E1",
  "Ctl-Male"    = "#415CAA",
  
  "EtOH-Female" = "#F6B284",
  "EtOH-Male"   = "#F26522",
  
  "Ctl"         = "#7A7A7A",  
  "EtOH"        = "#000000"
)


eae_data <- read_csv("../data/Extended_EAE_data.csv")

eae_data <- eae_data %>%
  mutate(pattern = paste(Treatment, Sex, sep = "-"))

comparison.plot("EtOH", "Ctl", "FigureS1_part1")
comparison.plot("EtOH-Male", "Ctl-Male", "FigureS1_part2")
comparison.plot("EtOH-Female", "Ctl-Female", "FigureS1_part3")
comparison.plot("Ctl-Male", "Ctl-Female", "FigureS1_part4")
comparison.plot("EtOH-Male", "EtOH-Female", "FigureS1_part5")
```

### Supplementary Figure 1B
```{r, fig.width = 6.5, fig.height = 4.5}
weights <- read_csv("../data/Mice_weights.csv")

weights_long <- weights %>%
  pivot_longer(cols = c(-Mouse_ID, -Treatment, -Sex), names_to = "Day", values_to = "Weight") %>%
  mutate(
    Day = as.numeric(Day),
    Group = factor(
      paste(Treatment, Sex, sep = "-"),
      levels = c("Ctl-Male", "EtOH-Male", "Ctl-Female", "EtOH-Female")
    )
  )

weight_summary <- weights_long %>%
  group_by(Group, Day) %>%
  summarise(
    mean_weight = mean(Weight, na.rm = TRUE),
    sem = standard_error(Weight),
    .groups = "drop"
  )

p_weight <- ggplot(weight_summary,
                   aes(Day, mean_weight, color = Group, group = Group)) +
  annotate("rect", xmin = 11, xmax = 17, ymin = -Inf, ymax = Inf,
           alpha = 0.1, fill = "grey") +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = mean_weight - sem,
                    ymax = mean_weight + sem),
                width = 0.3) +
  scale_color_manual(values = c(
    "Ctl-Male"="#415CAA","Ctl-Female"="#9BB9E1",
    "EtOH-Male"="#F26522","EtOH-Female"="#F6B284")) +
  labs(x = "Days post-induction", y = "Body weight (g)")
p_weight
ggsave("../output/FigureS1B.pdf", width = 6.5, height = 4.5)
```


## Supplementary Figure 2
```{r, fig.width = 5.5, fig.height = 2.5}
## data is the loaded Flow data from Figure 6
ggs1 <- data %>%
  pivot_longer(c(-`mouse_id`, -Group)) %>%
  filter(name %in% c("CD4+/Q1: CD69+ OX40-", "CD4+/Q2: CD69+ OX40+",
                     "CD4+/Q3: CD69- OX40+", "CD4+/Q4: CD69- OX40-")) %>%
  mutate(name = gsub(".*: ", "", name)) %>%
  mutate(name = factor(name, c("CD69- OX40-", "CD69+ OX40-",
                               "CD69+ OX40+", "CD69- OX40+"))) %>%
  ggplot(aes(x=Group, y= value, color = Group)) +
  geom_boxplot(outlier.shape = NA) +
  ggbeeswarm::geom_quasirandom() +
  theme_classic() +
  facet_wrap(~name, scales = "free", nrow = 1) +
  ylab("Percent of CD4+ T cells (%)")  +
  stat_compare_means(comparisons = list(c("EtOH", "Ctl")),
                     method = "wilcox.test") +
  scale_color_manual(values = color_pal)+
  theme(axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        strip.background = element_blank(),
        legend.position = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) 
ggs1
ggsave(filename = "FigureS2B_part1.pdf", path = "../output", width = 5.5, height = 2.5)

ggs1_2 <- data %>%
  pivot_longer(c(-`mouse_id`, -Group)) %>%
  filter(name %in% c("CD4+/Q1: OX40+ PD1-", "CD4+/Q2: OX40+ PD1+",
                     "CD4+/Q3: OX40- PD1+", "CD4+/Q4: OX40- PD1-")) %>%
  mutate(name = gsub(".*: ", "", name)) %>%
  mutate(name = factor(name, c("OX40- PD1-", "OX40+ PD1-",
                               "OX40+ PD1+", "OX40- PD1+"))) %>%
  ggplot(aes(x=Group, y= value, color = Group)) +
  geom_boxplot(outlier.shape = NA) +
  ggbeeswarm::geom_quasirandom() +
  theme_classic() +
  facet_wrap(~name, scales = "free", nrow = 1) +
  ylab("Percent of CD4+ T cells (%)")  +
  stat_compare_means(comparisons = list(c("EtOH", "Ctl")),
                     method = "wilcox.test") +
  scale_color_manual(values = color_pal)+
  theme(axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        strip.background = element_blank(),
        legend.position = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) 
ggs1_2
ggsave(filename = "FigureS2B_part2.pdf", path = "../output", width = 5.5, height = 2.5)

```

## SessionInfo and Package Versions
```{r}
sessionInfo()
```

R version 4.5.0 (2025-04-11 ucrt)
Platform: x86_64-w64-mingw32/x64
Running under: Windows 11 x64 (build 26100)

Matrix products: default
  LAPACK version 3.12.1

locale:
[1] LC_COLLATE=English_United States.utf8 
[2] LC_CTYPE=English_United States.utf8   
[3] LC_MONETARY=English_United States.utf8
[4] LC_NUMERIC=C                          
[5] LC_TIME=English_United States.utf8    

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] patchwork_1.3.2  vegan_2.7-1      permute_0.9-7    phyloseq_1.52.0 
 [5] rstatix_0.7.2    sciplot_1.2-0    Hmisc_5.2-5      writexl_1.5.4   
 [9] ggbeeswarm_0.7.2 emmeans_1.11.1   readxl_1.4.5     ggpubr_0.6.0    
[13] lubridate_1.9.4  forcats_1.0.0    stringr_1.5.1    dplyr_1.1.4     
[17] purrr_1.0.4      readr_2.1.5      tidyr_1.3.1      tibble_3.2.1    
[21] ggplot2_3.5.2    tidyverse_2.0.0 

loaded via a namespace (and not attached):
 [1] gridExtra_2.3           rlang_1.1.6             magrittr_2.0.3         
 [4] ade4_1.7-23             compiler_4.5.0          mgcv_1.9-1             
 [7] systemfonts_1.2.3       vctrs_0.6.5             reshape2_1.4.4         
[10] pkgconfig_2.0.3         crayon_1.5.3            fastmap_1.2.0          
[13] backports_1.5.0         XVector_0.48.0          labeling_0.4.3         
[16] rmarkdown_2.29          tzdb_0.5.0              UCSC.utils_1.4.0       
[19] ragg_1.4.0              bit_4.6.0               xfun_0.52              
[22] GenomeInfoDb_1.44.3     jsonlite_2.0.0          biomformat_1.36.0      
[25] rhdf5filters_1.20.0     Rhdf5lib_1.30.0         broom_1.0.8            
[28] parallel_4.5.0          cluster_2.1.8.1         R6_2.6.1               
[31] stringi_1.8.7           RColorBrewer_1.1-3      car_3.1-3              
[34] rpart_4.1.24            cellranger_1.1.0        estimability_1.5.1     
[37] Rcpp_1.0.14             iterators_1.0.14        knitr_1.50             
[40] base64enc_0.1-3         IRanges_2.42.0          splines_4.5.0          
[43] Matrix_1.7-3            nnet_7.3-20             igraph_2.1.4           
[46] timechange_0.3.0        tidyselect_1.2.1        rstudioapi_0.17.1      
[49] abind_1.4-8             yaml_2.3.10             codetools_0.2-20       
[52] lattice_0.22-6          plyr_1.8.9              Biobase_2.68.0         
[55] withr_3.0.2             evaluate_1.0.5          foreign_0.8-90         
[58] survival_3.8-3          Biostrings_2.76.0       pillar_1.10.2          
[61] carData_3.0-5           checkmate_2.3.3         foreach_1.5.2          
[64] stats4_4.5.0            generics_0.1.4          vroom_1.6.5            
[67] S4Vectors_0.46.0        hms_1.1.3               scales_1.4.0           
[70] xtable_1.8-4            glue_1.8.0              tools_4.5.0            
[73] data.table_1.17.2       ggsignif_0.6.4          mvtnorm_1.3-3          
[76] cowplot_1.1.3           rhdf5_2.52.1            grid_4.5.0             
[79] ape_5.8-1               colorspace_2.1-1        nlme_3.1-168           
[82] GenomeInfoDbData_1.2.14 beeswarm_0.4.0          htmlTable_2.4.3        
[85] vipor_0.4.7             Formula_1.2-5           cli_3.6.5              
[88] textshaping_1.0.1       gtable_0.3.6            digest_0.6.37          
[91] BiocGenerics_0.54.0     htmlwidgets_1.6.4       farver_2.1.2           
[94] htmltools_0.5.8.1       multtest_2.64.0         lifecycle_1.0.4        
[97] httr_1.4.7              bit64_4.6.0-1           MASS_7.3-65  